---
title: "Engineering Serie A Data"
author: "Ravi Solter"
date: "February 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = F,message = F)

library(dplyr)
library(tidyr)
library(zoo)

load(file = "Data/seriea_2018_19.rdata")
load(file="Data/archive_serie.rdata")
raw_scraped <- rbind(archive,seriea_2018_19)

```

****

This markdown documents the different datasets created from the scraped Serie A data. These include:

- **raw_scraped** - raw data with *season*, *round*, and *game_id* variables added.

- **team_records_df** - weekly result data for each match, including variables measuring form, points gained, overall season records, and rankings for each team. 

- **DF** - Design matrix for modeling on result where each row represents home and away stats for as single match. Except for form, no variables have been modified.

- **DF_trailing** - Design matrix for modeling on result. Similar to **DF**, but all predictors have been replaced with trailing averages from the previous five matches.

- **DF_completion** - Design matrix where future matches have their predictor variables estimated using matrix completion.


****

### Data

Scraped from the Serie A website, [http://www.legaseriea.it/en](http://www.legaseriea.it/en), the following data are collected on each team's performance in a given Serie A match:

- Season 
- Round
- Home team
- Away team
- Goals scored
- Goals conceded
- Saves
- Penalties
- Shots
- Shots on and off target
- Shots on and off target from freekicks
- Shots within the box
- Shots from set pieces
- Fouls
- Scoring chances
- Offsides
- Corners
- Yellow cards
- Red cards
- Balls lost
- Balls won
- Attacks down the middle
- Attacks down the left
- Attacks down the right

The following features are added:

- **Game ID**: A unique identifier for each match played in Serie A from the 2015-16 season onwards.

- **Team Form**: Form is represented by the proportion of points gained vs those possilbe for the previous five matches. A team with the following record: (W,D,W,D,L) would have collected 8 points from a possible and a form of 0.533. Home and away form are calculated in a similar fashion. Note that the first four matches of each season, teams' form is imputed with the their previous season's overall form.

- **Team Ranking**: Team's rankings within the league is based upon their form in the previous 38 matches. The form here is calcualted teh same as Team form.


```{r import,echo=F}
source(file="Processing_and_Engineering//engineering_seriea_design_matrix.R")
source(file="Processing_and_Engineering//engineering_seriea_design_matrix trailing.R")

```

****

### Datasets



**team_records_df**:

- *This dataframe includes records from 2015-16 season onwards*

- *There are two rows per match*


```{r team_records, echo=F}
glimpse(team_records_df)
```

****

**DF**:

- *This dataframe includes records from 2016-17 season onwards*

- *There is one row per match*

```{r DF, echo=F}
glimpse(DF)
```


****
**DF_trailing**:

- *This dataframe includes records from 2016-17 season onwards*

- *There are two rows per match with each match/row duplicated for both teams' perspective*

```{r DF_trailing, echo=F}
glimpse(DF_trailing)
```

